Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-18 paper_jasa.tex
comments colour removed

\begin{figure*}[hbtp]
   \centering
      \includegraphics[width=0.95\textwidth]{p_val_log_counts-a.pdf}
       \includegraphics[width=0.95\textwidth]{p_val_log_counts-b.pdf}
  %      \includegraphics[width=0.95\textwidth]{p_val_log_counts.pdf}
       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the sixty (60) individual lineups for experiment 1. \blue{Red indicates the} plot with the lowest $p$-value, and 
       \blue{blue indicates the} actual data plot. %, when it is different from that with the lowest $p$-value. 
       The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.
       %Rows correspond to $\beta$, smallest to largest values from top to bottom. Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate. Empty cells indicate no lineup for that combination of parameters. Highest counts tend to be the plot in the lineup having the lowest $p$-value.
       }
       \label{fig:P-val_log}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
%       \includegraphics[width=0.95\textwidth]{p_val_log_counts2.pdf}
       \includegraphics[width=0.75\textwidth]{p_val_log_counts2-a.pdf}       
       \includegraphics[width=0.75\textwidth]{p_val_log_counts2-b.pdf}
       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the seventy (70) individual lineups for experiment 2. \blue{Red indicates the} plot with the lowest $p$-value. %, and \blue{blue indicates} the actual data plot, when it is different from that with the lowest $p$-value. 
       %Rows correspond to $\beta$, smallest to largest values from top to bottom. Columns correspond to other experimental treatments, sample size $(n)$, error standard deviation $(\sigma)$, and replicate. 
%       
Columns are ordered according to effect size, with rows showing replicates of the same parameter combination on top of each other.
       Empty cells indicate combination of parameters with fewer than five replicates. 
       Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.
       }
       \label{fig:P-val_log2}
\end{figure*}

@@

We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2}). What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure. If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot. Figure \ref{fig:pval_difference} investigates this. The difference between the $p$-value of the actual data is compared with the lowest from the null plots. This is plotted horizontally, and the proportion correct is plotted vertically. Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots. In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, but only just. This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups. Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure. Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects' as one force choose the same plot. When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).

>>>>

\begin{figure*}[hbtp]
   \centering
      \includegraphics[width=0.95\textwidth]{p_val_log_counts-a.pdf}
       \includegraphics[width=0.95\textwidth]{p_val_log_counts-b.pdf}
       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 60 individual lineups for experiment 1. Red indicates the plot with the lowest $p$-value, and 
       blue indicates the actual data plot, when it is different from that with the lowest $p$-value. 
       The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.
       }
       \label{fig:P-val_log}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \includegraphics[width=0.75\textwidth]{p_val_log_counts2-a.pdf}       
       \includegraphics[width=0.75\textwidth]{p_val_log_counts2-b.pdf}
       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 70 individual lineups for experiment 2. Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value. 
Columns are ordered according to effect size, with rows showing replicates of the same parameter combination on top of each other.
       Empty cells indicate combination of parameters with fewer than five replicates. 
       Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.
       }
       \label{fig:P-val_log2}
\end{figure*}

@@

We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2}). What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure. If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot. Figure \ref{fig:pval_difference} investigates this. The difference between the $p$-value of the actual data is compared with the lowest from the null plots. This is plotted horizontally, and the proportion correct is plotted vertically. Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots. In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, but only just. This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups. Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure. Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot. When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).
